---
title: "Tracker_Data"
author: "Dionis Anderegg"
date: "16 Mai 2021"
output: html_document
---

```{r setup, include=FALSE}
# Load Packages
library(tidyverse)
library(sf)
library(lubridate)
library(ggmap)
library(zoo)
```

### Activity Tracker Data - How hight difference affects running speed

This report shows how hight difference affects the running speed of a person using acitivty data of a sports
tracker. 

## Read and explore data

```{r, echo=TRUE}
# Which layer needed
st_layers("activity_6531531145.gpx") # track_points needed

# Read data:
p = read_sf("activity_6531531145.gpx", layer = "track_points")

# Convert to CH1903+ LV95
p <- st_transform(p, crs = 2056)

# When was the measurement?
summary(p$time)
plot(p$time)

```

The duration of the activity was `r max(p$time) - min(p$time)` hours.

## Calculate timelag, steplength, speed and hight difference
Time lag is measured in seconds, euclidian steplength in meters, speed in km/h and hight difference in meters.

# Timelag
```{r, echo=TRUE}
# Timelag
p$timelag <- as.integer(difftime(lead(p$time), p$time, units = "secs"))
summary(p$timelag)
ggplot(p, aes(x = timelag)) +
  geom_histogram(binwidth = 1) +
  theme_bw()
```

Time difference is between `r min(p$timelag, na.rm = TRUE)` and `r max(p$timelag, na.rm = TRUE)` seconds. 

# Euclidian distance
```{r, echo=TRUE}
# Euclidian distance
options(digits=2)
cords <- st_coordinates(p$geometry)

p$E <- cords[,1]
p$N <- cords[,2]

p$steplength <- sqrt(
  (p$E - lead(p$E))^2 + (p$N - lead(p$N))^2
)

```

The total distance tracked was `r as.integer(sum(p$steplength, na.rm = TRUE))` meters. 

# Speed
```{r, echo=TRUE}
# Speed
p$speed_ms <- (p$steplength / p$timelag)
p$speed_kmh <- (p$speed_ms * 3.6)


p$speed_kmh_smooth_03 <- rollmean(p$speed_kmh, k = 3, fill = NA, allign = "left")
p$speed_kmh_smooth_05 <- rollmean(p$speed_kmh, k = 5, fill = NA, allign = "left")
p$speed_kmh_smooth_10 <- rollmean(p$speed_kmh, k = 10, fill = NA, allign = "left")

p %>%
  gather(k, Geschwindigkeit, speed_kmh : speed_kmh_smooth_10) %>%
  ggplot(aes(time, Geschwindigkeit, col = k)) +
  geom_line() +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(title = "Speed diagram\n", x = "\ntime [hh:mm]", y = "speed [kmh]\n" )
```

Speed strongly depends on the used k-value in rollmeans. We use k = 5 for the following results.

```{r, echo=TRUE}
p %>%
  ggplot(aes(time, speed_kmh_smooth_03)) +
  geom_line(col = "blue") +
  theme_bw() +
  labs(title = "Speed diagram\n", x = "\ntime [hh:mm]", y = "speed [kmh]\n" )
```

# Hight difference
```{r, echo=TRUE}
p$hightdiff <- lead(p$ele) - p$ele
ggplot(p, aes(time, hightdiff)) +
  geom_line()
```

Hight difference of every timestep seems to be not accurate => See Plot!
We have to analyse bigger timeframes!

```{r, echo=TRUE}
p$hightdiff <- ifelse(abs(p$hightdiff) > 2, NA, p$hightdiff)
```

# Show the Track on a map
```{r, echo=FALSE}
# Track
ggplot(p, aes(E, N, col = speed_kmh_smooth_03)) +
  geom_path() +
  geom_point() + 
  scale_color_gradient(low = "red", high = "yellow")

a <- st_transform(p, crs = 4326)

acords <- cords <- st_coordinates(a$geometry)
a$E <- acords[,1]
a$N <- acords[,2]


# Show the track on a map:
myLocation<-c(min(a$E)-.01, min(a$N)-.01, max(a$E)+.01, max(a$N)+.01)
myMap <- get_stamenmap(bbox=myLocation, maptype="terrain", crop=TRUE, zoom = 13)
ggmap(myMap) +
  geom_point(data = a, aes(E,N, col = speed_kmh_smooth_03)) +
  labs(title = paste(date(a$time), "Trailrunning"),
       x = "\n째E", y = "째N\n") +
    scale_color_gradient(low = "red", high = "yellow")

ggmap(myMap) +
  geom_point(data = a, aes(E,N, col = hightdiff)) +
  labs(title = paste(date(a$time), "Trailrunning"),
       x = "\n째E", y = "째N\n") +
    scale_color_gradient(low = "yellow", high = "red")


```

As the two figures above show, the speed seems to be strongly correlated with the hight difference!

## Correlation of speed and hight difference
# Group the relevant data into intervals of 1 minute
```{r, echo=TRUE}
mins <- a %>%
  st_drop_geometry() %>%
  select(time, steplength, hightdiff) %>%
  group_by("time" = cut(time, "1 min")) %>%
  summarise(steplength = sum(steplength, na.rm = T),
            hightdiff_up = sum(hightdiff > 0),
            hightdiff_down = sum(hightdiff < 0),
            speed = steplength / 60 * 3.6)

mins$speed <- ifelse(mins$speed < 3, NA, mins$speed)

sum(mins$hightdiff_up, na.rm = T)
sum(mins$hightdiff_down, na.rm = T)
  
```

# Show heightdifference vs speed
```{r, echo=FALSE}
mins_long <- gather(mins, updown, hightdiff, hightdiff_up : hightdiff_down)

ggplot(mins_long, aes(hightdiff, speed, col = updown)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(x = "\n hight difference", y = "speed\n")
```


# Calculate a simple linear model of speed and hight difference
```{r, echo=FALSE}
cor(mins$speed, mins$hightdiff_up, use = "pairwise.complete.obs")
lm_up <- lm(mins$speed~mins$hightdiff_up)
summary(lm_up)
```

Interpretation of the results: The speed is strongly correlated with the hight difference (up)
with a correlation coefficient of `cor(mins$speed, mins$hightdiff_up, use = "pairwise.complete.obs")`
and an adjusted R-squared of 0.68. The gorrelation is highly significant with p < 0.001.
       